### 2019-05-06 fgc_table.tex
Autocommit (fgc)

\documentclass[hidelinks,10pt,letterpaper]{article}
\usepackage[top=0.75in,footskip=0.75in,bottom=0.75in]{geometry}
\usepackage{float}
\usepackage{changepage}
\usepackage[utf8]{inputenc}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage{cite}
\usepackage{nameref,hyperref}
\usepackage[right]{lineno}
\usepackage{microtype}
% \DisableLigatures[f]{encoding = *, family = * }
% \raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
%% \textheight 8.75in
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother
\date{}
\usepackage{lastpage,fancyhdr,graphicx}
\pagestyle{myheadings}
\pagestyle{fancy}

% \fancyhf{}
% \setlength{\headheight}{27.023pt}
% \rfoot{\thepage/\pageref{LastPage}}
% \renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
% \fancyheadoffset[L]{2.25in}
% \fancyfootoffset[L]{2.25in}

\usepackage{setspace}

\begin{document}
% \vspace*{0.35in}
\section*{Supplementary Table: Sample Characteristics}

\setcounter{table}{0}
\renewcommand\thetable{\Alph{table}}

%% small blurb here

\begin{table}[H]

\caption{FGC table}
\scalebox{0.9}{
\tiny{
\begin{tabular}{|l|rrrr|r|}
\hline
\bf{Country} & KE & ML & NG & SL & Total \\
\hline
\input{table_output.tex}
\hline
\end{tabular}
}
}
\\
\\
\textrm{Feed me}
\label{fgc_table}
\end{table}

\end{document}